---
title: "Getting Started with geographr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with geographr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(geographr)
```

## Loading the Data

`load_data()` loads the geopolitical dataset containing country-level economic,
health, and regime classification data.

```{r}
dat <- load_data()
head(dat)
```

## Filtering by Year

`filter_year()` subsets the data to a specific year range. By default it keeps
2000–2020.

```{r}
recent <- filter_year(dat, start_year = 2010, end_year = 2015)
range(recent$year)
```

## Summarizing GDP by Regime Type

`summarize_gdp_by_regime()` computes the mean, median, and count of GDP per
capita grouped by regime category. You can optionally filter to a single year.

```{r}
summarize_gdp_by_regime(dat)
```

```{r}
summarize_gdp_by_regime(dat, year = 2015)
```

## Summarizing GDP by Democracy Status

`summarize_gdp_by_democracy()` computes mean GDP per capita for democracies
vs non-democracies by year. Uses `data.table` under the hood for performance.

```{r}
dem_summary <- summarize_gdp_by_democracy(dat)
head(dem_summary)
```

## Plotting GDP by Regime Type

`plot_regime_gdp()` creates a faceted area plot showing mean GDP per capita
over time for each regime group.

```{r, fig.width=8, fig.height=5}
plot_regime_gdp(dat, start_year = 2000, end_year = 2020)
```

## Plotting Presidential Tenure by Regime

`plot_tenure_ridge()` creates a density ridge plot showing the distribution
of presidential tenure for selected regime types.

```{r, fig.width=8, fig.height=5}
plot_tenure_ridge(
  regimes = c("Presidential democracy", "Military dictatorship"),
  data = dat
)
```
  
